---
title: '`r rstudioapi::getSourceEditorContext()$path %>% basename() %>% gsub(pattern = "\\.Rmd",replacement = "")`' 
author: "Avishai Wizel"
date: '`r Sys.time()`'
output: 
  html_notebook: 
    code_folding: hide
    toc: yes
    toc_collapse: yes
    toc_float: 
      collapsed: FALSE
    number_sections: true
    toc_depth: 1
---



# Functions


```{r warning=FALSE}

library(stringi)
library(reticulate)
source_from_github(repositoy = "DEG_functions",version = "0.2.45")
source_from_github(repositoy = "cNMF_functions",version = "0.4.0",script_name = "cnmf_functions_V3.R")
source_from_github(repositoy = "sc_general_functions",version = "0.1.0",script_name = "functions.R")

```


# Data

```{r}
genesets <- msigdb_download("Homo sapiens",category="H") %>% append( msigdb_download("Homo sapiens",category="C2",subcategory = "CP:KEGG"))
genesets[["HIF_targets"]] = hif_targets
```

```{r fig.height=5, fig.width=12}
DimPlot(xeno,group.by = "orig.ident")+ DimPlot(xeno,group.by = "treatment",shuffle = T)
```


```{python}
from cnmf import cNMF
import pickle
f = open('./Data/cnmf/cnmf_objects/models_2Kvargenes_all_K_cnmf_obj.pckl', 'rb')
cnmf_obj = pickle.load(f)
f.close()
```

# K selection plot
```{r fig.height=2, fig.width=2}
plot_path = paste0("/sci/labs/yotamd/lab_share/avishai.wizel/R_projects/EGFR/Data/cNMF/cNMF_models_Varnorm_Harmony_2Kvargenes_all_K/cNMF_models_Varnorm_Harmony_2Kvargenes_all_K.k_selection.png")
knitr::include_graphics(plot_path)
```

```{python}
k = 5
density_threshold = 0.1 
cnmf_obj.consensus(k=k, density_threshold=density_threshold,show_clustering=True)
usage_norm, gep_scores, _, _ = cnmf_obj.load_results(K=k, density_threshold=density_threshold)
```

```{r}
usage_norm5_xeno  = py$usage_norm
gep_scores5_xeno = py$gep_scores
```

```{r}
gep_scores = gep_scores5_xeno
usage_norm = usage_norm5_xeno
```

# NMF usage
```{r fig.height=8, fig.width=12, results='asis'}
  for (i in 1:ncol(usage_norm)) {
    metage_metadata = usage_norm %>% dplyr::select(i)
    xeno = AddMetaData(object = xeno,metadata = metage_metadata,col.name = paste0("gep",i))
  }
  
  FeaturePlot(object = xeno,features = paste0("gep",1:ncol(usage_norm)),ncol = 3)
```
# Programs GSEA {.tabset}

```{r results='asis'}
  for (col in seq_along(gep_scores)) {
     ranked_vec = gep_scores[,col] %>% setNames(rownames(gep_scores)) %>% sort(decreasing = TRUE) 
     hyp_obj <- hypeR_fgsea(ranked_vec, genesets)
       print_tab(hyp_dots(hyp_obj,title = paste("program",col))+ aes(size=nes),title = paste0("gep",col))
  }
```

```{r}
programs_main_pathways = list(gep1 = "HALLMARK_INTERFERON_ALPHA_RESPONSE", gep2 = c("HALLMARK_TNFA_SIGNALING_VIA_NFKB","KEGG_ASTHMA","KEGG_TOLL_LIKE_RECEPTOR_SIGNALING_PATHWAY"),gep3 = c("HALLMARK_HYPOXIA","HIF_targets"),gep4 = "HALLMARK_E2F_TARGETS")
```

```{r}
xeno = FindVariableFeatures(object = xeno,nfeatures = 2000)
xeno_vargenes = VariableFeatures(object = xeno)

xeno_expression = FetchData(object = xeno,vars = xeno_vargenes,slot='counts')
all_0_genes = colnames(xeno_expression)[colSums(xeno_expression==0, na.rm=TRUE)==nrow(xeno_expression)] #delete rows that have all 0
xeno_vargenes = xeno_vargenes[!xeno_vargenes %in% all_0_genes]

```


# calculate score for Xeno
```{python}
import numpy as np
import scanpy as sc
xeno_expression = r.xeno_expression
xeno_vargenes = r.xeno_vargenes
tpm =  compute_tpm(xeno_expression)
usage_by_calc = get_usage_from_score(counts=xeno_expression,tpm=tpm,genes=xeno_vargenes, cnmf_obj=cnmf_obj,k=k)
```

```{r}
xeno_5_metagenes = py$usage_by_calc
```

```{r}
all_metagenes = xeno_5_metagenes
colnames(all_metagenes) = c("IFNa","immune_response", "hypoxia","cell_cycle","unknown")

```

# programs expression
```{r echo=TRUE, fig.height=7, fig.width=12, results='asis'}

#add each metagene to metadata
for (i  in 1:ncol(all_metagenes)) {
  metagene_metadata = all_metagenes[,i,drop=F]
  xeno = AddMetaData(object = xeno,metadata = metagene_metadata,col.name = names(all_metagenes)[i])
}

FeaturePlot(object = xeno,features = colnames(all_metagenes),ncol = 3)


```
# Programs dotplot
```{r fig.width=8}
DotPlot(object = xeno, features =  colnames(all_metagenes),group.by  = 'treatment',scale = T)+
  guides(size = guide_legend(title = "Cells expressing (%)"),color = guide_colorbar(title = "Average Score"))
```




```{r}
#rename unknown to resistant program
xeno$resistant_program <- xeno$unknown
xeno$unknown <- NULL
all_metagenes = all_metagenes %>% dplyr::rename(resistant_program=unknown)
```

# NMF programs regulation  {.tabset}

```{r echo=TRUE,  results='asis'}
metagenes_mean_compare(dataset = xeno,time.point_var = "treatment",prefix = "model",patient.ident_var = "orig.ident",pre_on = c("NT","OSI"),test = "wilcox.test",programs = colnames(all_metagenes)[1:4],without_split = F)
```


# LE genes programs UMAP  {.tabset}


```{r results='asis'}
  for (col in seq_along(gep_scores[1:4])) {
     ranked_vec = gep_scores[,col] %>% setNames(rownames(gep_scores)) %>% sort(decreasing = TRUE) 
     hyp_obj <- hypeR_fgsea(ranked_vec, genesets)
     for (pathway in programs_main_pathways[[col]]) {
        le_genes =  hyp_obj$data %>% filter(label == pathway) %>% pull("le") %>% strsplit(",") %>% unlist()
        scoresAndIndices = getPathwayScores(dataMatrix = xeno@assays$RNA@data,pathwayGenes = le_genes)
        pathway_name = paste0(pathway,"_le")
        xeno=AddMetaData(xeno,scoresAndIndices$pathwayScores,col.name = pathway_name)
        print_tab(FeaturePlot(object = xeno,features = pathway_name),title = pathway_name)
     }
  }
```


# LE genes programs regulation  {.tabset}

```{r  results='asis'}
metagenes_mean_compare(dataset = xeno,time.point_var = "treatment",prefix = "model",patient.ident_var = "orig.ident",pre_on = c("NT","OSI"),test = "wilcox.test",programs = programs_main_pathways %>% unlist() %>% paste0("_le"),without_split = F)
```
```{r}
col=2
ranked_vec = gep_scores[, col] %>% setNames(rownames(gep_scores)) %>% sort(decreasing = TRUE)
print (paste("running gep",col))
hyp_obj <-hypeR_fgsea(ranked_vec, genesets_go, up_only = T)

print(hyp_dots(hyp_obj, title = paste("program", col), abrv = 70) + aes(size =nes))
  
```

# Top program 2 genes expression correlation
```{r}
top_ot = gep_scores [order(gep_scores [,2],decreasing = T),2,drop = F]%>% head(200) %>% rownames()

num_of_clusters = 7
annotation = plot_genes_cor(dataset = xeno,num_of_clusters = num_of_clusters,geneIds = top_ot,height = 3)

```

#  program 2 all clusters expression {.tabset}
```{r results='asis',fig.width=14}
for (chosen_clusters in 1:num_of_clusters) {
  chosen_genes = annotation %>% dplyr::filter(cluster == chosen_clusters) %>% rownames() #take relevant genes
  # print(chosen_genes)
  hyp_obj <- hypeR(chosen_genes, genesets, test = "hypergeometric", fdr=1, plotting=F,background = rownames(xeno_5_gep_scores))

   scoresAndIndices <- getPathwayScores(xeno@assays$RNA@data, chosen_genes)
  xeno=AddMetaData(xeno,scoresAndIndices$pathwayScores,paste0("cluster",chosen_clusters))

  
  print_tab(plt = 
              hyp_dots(hyp_obj,size_by = "none",title = paste0("cluster",chosen_clusters))+
              FeaturePlot(object = xeno,features = paste0("cluster",chosen_clusters)),
            title = chosen_clusters)
}


```

# Correlation of clusters
```{r}
for (chosen_clusters in 1:num_of_clusters) {
  
  cor_res = cor(xeno$TNFa,xeno[[paste0("cluster",chosen_clusters)]])
print(paste("correlation of TNFa program to", paste0("cluster",chosen_clusters),":", cor_res))

}
```


```{r}
clusters_idents = c("cluster1", "KEGG_OXIDATIVE_PHOSPHORYLATION","HALLMARK_TNFA_SIGNALING_VIA_NFKB","KEGG_ANTIGEN_PROCESSING_AND_PRESENTATION","HALLMARK_P53_PATHWAY","cluster6","cluster7")
```


#  program 2 intersected genes



```{r}
programs_of_cluster = c()
for (chosen_clusters in 1:num_of_clusters) {
  chosen_genes = annotation[["myannotation"]] %>% dplyr::filter(cluster == chosen_clusters) %>% rownames() #take relevant genes
  pathway_name = clusters_idents[chosen_clusters]
  if (!startsWith(x = pathway_name,prefix = "cluster")){
      chosen_genes  = (chosen_genes) %>% intersect(genesets[[pathway_name]])
      pathway_name = paste0(pathway_name,"_cluster")
  }
  programs_of_cluster = c(programs_of_cluster,pathway_name)
  print(pathway_name)
  print(chosen_genes)
  cat("\n")
  scoresAndIndices <- getPathwayScores(xeno@assays$RNA@data, chosen_genes)
  xeno=AddMetaData(xeno,scoresAndIndices$pathwayScores,pathway_name)

}
```
#  program 2 intersected pathway regulation {.tabset}     

```{r  results='asis'}
metagenes_mean_compare(dataset = xeno,time.point_var = "treatment",prefix = "model",patient.ident_var = "orig.ident",pre_on = c("NT","OSI"),test = "wilcox.test",programs = programs_of_cluster,without_split = F)
```
#  program 2 significant plot  

```{r fig.height=12}
signf_plot_pre_vs_on<- function(dataset,programs,patient.ident_var,prefix,pre_on,test,time.point_var) {
    final_df = data.frame()
    for (metegene in programs) {
      genes_by_tp = FetchData(object = dataset,vars = metegene) %>% rowSums() %>% as.data.frame() #mean expression
      names(genes_by_tp)[1] = metegene
      genes_by_tp = cbind(genes_by_tp,FetchData(object = dataset,vars = c(patient.ident_var,time.point_var))) # add id and time points
      
      
      genes_by_tp_forPlot =  genes_by_tp %>% mutate(!!ensym(patient.ident_var) := paste(prefix,genes_by_tp[,patient.ident_var])) #add "model" before  each model/patient
      fm <- as.formula(paste(metegene, "~", time.point_var)) #make formula to plot
      
      #plot and split by patient:   
      stat.test = compare_means(formula = fm ,data = genes_by_tp_forPlot,method = test,group.by = patient.ident_var)%>% 
              dplyr::filter(group1 == pre_on[1] & group2 == pre_on[2])  #filter for pre vs on treatment only
      final_df = rbind(final_df,stat.test)
    }
    return(final_df)
}

undebug(signf_plot_pre_vs_on)
final_df = signf_plot_pre_vs_on(dataset = xeno,time.point_var = "treatment",prefix = "model",patient.ident_var = "orig.ident",pre_on = c("NT","OSI"),test = "wilcox.test",programs = programs_of_cluster )
final_df = reshape2::dcast(final_df, orig.ident  ~.y.,value.var = "p.adj") %>% column_to_rownames("orig.ident")

sig_heatmap(all_patients_result = final_df,title = "ad")
```
# Top program 3 genes expression correlation
```{r}
top_hypoxia = gep_scores [order(gep_scores [,3],decreasing = T),2,drop = F]%>% head(200) %>% rownames()

num_of_clusters = 4
annotation = plot_genes_cor(dataset = xeno,hallmark_name = NULL,num_of_clusters = num_of_clusters,geneIds = top_hypoxia)

```
#  program 3 all clusters expression {.tabset}

```{r results='asis',fig.width=14}
for (chosen_clusters in 1:num_of_clusters) {
  chosen_genes = annotation[["myannotation"]] %>% dplyr::filter(cluster == chosen_clusters) %>% rownames() #take relevant genes
  # print(chosen_genes)
  hyp_obj <- hypeR(chosen_genes, genesets_env, test = "hypergeometric", fdr=1, plotting=F,background = rownames(xeno_5_gep_scores))

   scoresAndIndices <- getPathwayScores(xeno@assays$RNA@data, chosen_genes)
  xeno=AddMetaData(xeno,scoresAndIndices$pathwayScores,paste0("cluster",chosen_clusters))

  
  print_tab(plt = 
              hyp_dots(hyp_obj,size_by = "none",title = paste0("cluster",chosen_clusters))+
              FeaturePlot(object = xeno,features = paste0("cluster",chosen_clusters)),
            title = chosen_clusters)


}


```

# Correlation of clusters
```{r}
for (chosen_clusters in 1:num_of_clusters) {
  
  cor_res = cor(xeno$hypoxia,xeno[[paste0("cluster",chosen_clusters)]])
print(paste("correlation of hypoxia program to", paste0("cluster",chosen_clusters),":", cor_res))

}
```

```{r}
clusters_idents = c("HALLMARK_HYPOXIA", "HIF_targets","cluster3","cluster4")
```

```{r}
programs_of_cluster = c()
for (chosen_clusters in 1:num_of_clusters) {
  chosen_genes = annotation[["myannotation"]] %>% dplyr::filter(cluster == chosen_clusters) %>% rownames() #take relevant genes
  pathway_name = clusters_idents[chosen_clusters]
  if (!startsWith(x = pathway_name,prefix = "cluster")){
      chosen_genes  = (chosen_genes) %>% intersect(genesets[[pathway_name]])
      pathway_name = paste0(pathway_name,"_cluster")
  }
  programs_of_cluster = c(programs_of_cluster,pathway_name)
  print(pathway_name)
  print(chosen_genes)
  cat("\n")
  scoresAndIndices <- getPathwayScores(xeno@assays$RNA@data, chosen_genes)
  xeno=AddMetaData(xeno,scoresAndIndices$pathwayScores,pathway_name)

}
```
#  program 3 intersected pathway regulation {.tabset}     

```{r results='asis'}
metagenes_mean_compare(dataset = xeno,time.point_var = "treatment",prefix = "model",patient.ident_var = "orig.ident",pre_on = c("NT","OSI"),test = "wilcox.test",programs = programs_of_cluster,without_split = F)
```

# Top program 3 genes expression correlation

```{r}
top_cc = gep_scores [order(gep_scores [,4],decreasing = T),2,drop = F]%>% head(200) %>% rownames()

num_of_clusters = 4
annotation = plot_genes_cor(dataset = xeno,hallmark_name = NULL,num_of_clusters = num_of_clusters,geneIds = top_cc)

```
#  program 3 all clusters expression {.tabset}

```{r results='asis',fig.width=14}
for (chosen_clusters in 1:num_of_clusters) {
  chosen_genes = annotation[["myannotation"]] %>% dplyr::filter(cluster == chosen_clusters) %>% rownames() #take relevant genes
  # print(chosen_genes)
  hyp_obj <- hypeR(chosen_genes, genesets_env, test = "hypergeometric", fdr=1, plotting=F,background = rownames(xeno_5_gep_scores))

   scoresAndIndices <- getPathwayScores(xeno@assays$RNA@data, chosen_genes)
  xeno=AddMetaData(xeno,scoresAndIndices$pathwayScores,paste0("cluster",chosen_clusters))

  
  print_tab(plt = 
              hyp_dots(hyp_obj,size_by = "none",title = paste0("cluster",chosen_clusters))+
              FeaturePlot(object = xeno,features = paste0("cluster",chosen_clusters)),
            title = chosen_clusters)
  
}


```

<script src="https://hypothes.is/embed.js" async></script>



