---
title: '`r rstudioapi::getSourceEditorContext()$path %>% basename() %>% gsub(pattern = "\\.Rmd",replacement = "")`' 
author: "Avishai Wizel"
date: '`r Sys.time()`'
output: 
  html_notebook: 
    code_folding: hide
    toc: yes
    toc_collapse: yes
    toc_float: 
      collapsed: FALSE
    number_sections: true
    toc_depth: 1
---



# Functions
```{r warning=FALSE}

library(stringi)
library(reticulate)
source_from_github(repositoy = "DEG_functions",version = "0.2.24")
source_from_github(repositoy = "cNMF_functions",version = "0.3.91",script_name = "cnmf_function_Harmony.R")

no_neg <- function(x) {
  x = x + abs(min(x))
  x
}

sum_2_one <- function(x) {
  x =x/sum(x)
  x
}

```




# Data
```{r}
xeno = readRDS("./Data/10x_xeno_1000.Rds")
lung = readRDS("./Data/lung_cancercells_withTP_onlyPatients.rds")
lung_patients = lung$patient.ident %>% unique() %>% as.character()
lung_patients_filtered = lung_patients[!(lung_patients %in% c("X1055new","X1099"))] # remove patients with less than 100 malignant cells
lung = subset(x = lung,subset = patient.ident %in% lung_patients_filtered)
```


# Models 2K vargenes 

```{python}
import pickle
from cnmf import cNMF
f = open('./Data/cnmf/cnmf_objects/models_2Kvargenes_cnmf_obj.pckl', 'rb')
cnmf_obj = pickle.load(f)
f.close()
```

# K selection plot
```{r fig.height=2, fig.width=2}
plot_path = paste0("/sci/labs/yotamd/lab_share/avishai.wizel/R_projects/EGFR/Data/cnmf/cNMF_models_Varnorm_Harmony_2Kvargenes/cNMF_models_Varnorm_Harmony_2Kvargenes.k_selection.png")
knitr::include_graphics(plot_path)
```



# gep scores for all NMF k's {.tabset}
```{python}
density_threshold = 0.1
usage_norm3, gep_scores3, _, _ = cnmf_obj.load_results(K=3, density_threshold=density_threshold)
usage_norm4, gep_scores4, _, _ = cnmf_obj.load_results(K=4, density_threshold=density_threshold)
usage_norm5, gep_scores5, _, _ = cnmf_obj.load_results(K=5, density_threshold=density_threshold)
```

```{r}
gep_scores3 = py$gep_scores3
gep_scores4 = py$gep_scores4
gep_scores5 = py$gep_scores5

usage_norm3 = py$usage_norm3
usage_norm4 = py$usage_norm4
usage_norm5 = py$usage_norm5

```

```{r fig.height=7, fig.width=10, results='asis'}
all_usage_norm = list(usage_norm3 = usage_norm3, usage_norm4 = usage_norm4, usage_norm5 = usage_norm5)
for (usage_num in seq_along(all_usage_norm)) {
  usage = all_usage_norm[[usage_num]]
  #add each metagene to metadata
  for (i in 1:ncol(usage)) {
    metage_metadata = usage %>% dplyr::select(i)
    xeno = AddMetaData(object = xeno,metadata = metage_metadata,col.name = paste0("gep",i))
  }
  if (usage_num==1) {
      print_tab(FeaturePlot(object = xeno,features = paste0("gep",1:ncol(usage)),ncol = 2),title = usage_num)

  }else{ print_tab(FeaturePlot(object = xeno,features = paste0("gep",1:ncol(usage)),ncol = 3),title = usage_num)}
}
```
# all NMF k's  enrichment{.tabset}


```{r results='asis'}

all_gep_scores =  list(gep_scores3 = gep_scores3, gep_scores4 = gep_scores4, gep_scores5 = gep_scores5)

genesets <- msigdb_download("Homo sapiens",category="H") %>% append( msigdb_download("Homo sapiens",category="C2",subcategory = "CP:KEGG"))
genesets[["HIF_targets"]] = hif_targets
genesets_env = gsets$new(genesets, name="my genesets", version="v1.0")

for (gep_num in seq_along(all_gep_scores)) {
  gep = all_gep_scores[[gep_num]]
  ranked_list = list()
  for (col in (gep)) {
   lst = col %>% setNames(rownames(gep)) %>% sort(decreasing = TRUE) 
   ranked_list  %<>% append(list(lst))
  }
  names(ranked_list) = paste0("gep",1:ncol(gep))
  
  
  
  hyp_obj <- hypeR(ranked_list, genesets_env, test="kstest", fdr=0.05, plotting=F,background = rownames(gep_scores3))
  print_tab(
    hyp_dots(hyp_obj,size_by = "significance",abrv = 100,merge = T)+
      scale_color_continuous(low = "red", high = "black", guide = guide_colorbar(reverse = TRUE))
    ,title = gep_num)
  
}

```






```{r}
  for (col in seq_along(gep_scores3)) {
     ranked_vec = gep_scores3[,col] %>% setNames(rownames(gep_scores3)) %>% sort(decreasing = TRUE) 
     hyp_obj <- fgsea.wrapper(ranked_vec, genesets)
    # hyp_list[[paste0("gep",col)]] = hyp_obj
       print(hyp_dots(hyp_obj))
  }
```


```{r}
col = 2
 ranked_vec = gep_scores3[,col] %>% setNames(rownames(gep_scores3)) %>% sort(decreasing = TRUE) 
 hyp_obj <- fgsea.wrapper(ranked_vec, genesets)
# hyp_list[[paste0("gep",col)]] = hyp_obj
 for (i in 1:5) {
  le_genes =  hyp_obj$data[i,"le",drop=T] %>% strsplit(",") %>% unlist()
  scoresAndIndices = getPathwayScores(dataMatrix = xeno@assays$RNA@data,pathwayGenes = le_genes)
  xeno=AddMetaData(xeno,scoresAndIndices$pathwayScores,paste0(hyp_obj$data[i,"label",drop=T],"_gsea"))
  print(FeaturePlot(object = xeno,features = paste0(hyp_obj$data[i,"label",drop=T],"_gsea")))
 }

  
```


# programs enrichment
```{r}
gep_scores = py$gep_scores3
usage_norm= py$usage_norm3
```

```{r}
# or:
gep_scores = readRDS("/sci/labs/yotamd/lab_share/avishai.wizel/R_projects/EGFR/Data/cnmf/harmony_models_gep_scores.rds")
```

```{r fig.height=6, fig.width=8}
names (gep_scores) = c("Hypoxia","TNFa","Cell_cycle")
plt_list = list()

for (program  in names (gep_scores)) {
 p = ggplot(gep_scores, aes(x=!!ensym(program))) +
  geom_density()+xlab(program)+
   geom_vline(
    aes(xintercept=sort(gep_scores[,program],TRUE)[200]  ,color="top200"),
          linetype="dashed", size=1)+
   geom_vline(
    aes(xintercept=sort(gep_scores[,program],TRUE)[100]  ,color="top100"),
          linetype="dashed", size=1)+
      geom_vline(
    aes(xintercept=sort(gep_scores[,program],TRUE)[50]  ,color="top50"),
          linetype="dashed", size=1)+
         geom_vline(
    aes(xintercept=sort(gep_scores[,program],TRUE)[150]  ,color="top150"),
          linetype="dashed", size=1)+
   scale_color_manual(name = "top n genes", values = c(top200 = "blue",top100 = "red",top150 = "yellow",top50 = "green"))
   plt_list[[program]] <- p

}
 
ggarrange(plotlist = plt_list)

```





# Programs hallmark enrichment

```{r fig.height=8, fig.width=8}
ntop = 150
plt_list = list()
hif_targets_set = data.frame(gs_name = "hif_targets",gene_symbol = hif_targets)

for (i in 1:ncol(gep_scores)) {
  top_genes = gep_scores  %>%  arrange(desc(gep_scores[i])) #sort by score a
  top = head(rownames(top_genes),ntop) #take top top_genes_num
  res = genes_vec_enrichment(genes = top,background = rownames(gep_scores),homer = T,title = 
                    i,silent = T,return_all = T,custom_pathways  = hif_targets_set)
   
  plt_list[[i]] = res$plt
}
gridExtra::grid.arrange(grobs = plt_list)
```

# program 2 all cp
```{r}
ntop = 150
i = 2 

  top_genes = gep_scores  %>%  arrange(desc(gep_scores[i])) #sort by score a
  top = head(rownames(top_genes),ntop) #take top top_genes_num
  res = genes_vec_enrichment(genes = top,background = rownames(gep_scores),homer = T,title = 
                    i,silent = T,return_all = T,custom_pathways  = canonical_pathways)
   
 print(res$plt) 


```

# Program 2 all msigdb
```{r}
ntop = 150
i = 2 

  top_genes = gep_scores  %>%  arrange(desc(gep_scores[i])) #sort by score a
  top = head(rownames(top_genes),ntop) #take top top_genes_num
  res = genes_vec_enrichment(genes = top,background = rownames(gep_scores),homer = T,title = 
                    i,silent = T,return_all = T,custom_pathways  = all_pathways)
   
 print(res$plt) 


```


```{r}
xeno = FindVariableFeatures(object = xeno,nfeatures = 2000)
xeno_vargenes = VariableFeatures(object = xeno)

xeno_expression = FetchData(object = xeno,vars = xeno_vargenes,slot='counts')
all_0_genes = colnames(xeno_expression)[colSums(xeno_expression==0, na.rm=TRUE)==nrow(xeno_expression)] #delete rows that have all 0
xeno_vargenes = xeno_vargenes[!xeno_vargenes %in% all_0_genes]

```



# Test with expr after harmony
```{python}
# import numpy as np
# import scanpy as sc
# 
# expr_after_harmony = sc.read_h5ad('./Data/cnmf/xeno_Harmony_NoNeg_2Kvargenes.h5ad').to_df()
# tpm =  compute_tpm(expr_after_harmony)
# cnmf_genes = expr_after_harmony.keys().to_list()
# usage_by_calc = get_usage_from_score(counts=expr_after_harmony,tpm=tpm,genes=cnmf_genes,cnmf_obj=cnmf_obj,k=3)
```

# Check if original cNMF score is like the calculated score
```{r}
usage_by_calc = py$usage_by_calc
usage_norm = py$usage_norm3
cor(usage_by_calc,usage_norm)
```

# calculate score for Xeno
```{python}
import numpy as np
import scanpy as sc
xeno_expression = r.xeno_expression
xeno_vargenes = r.xeno_vargenes
tpm =  compute_tpm(xeno_expression)
usage_by_calc = get_usage_from_score(counts=xeno_expression,tpm=tpm,genes=xeno_vargenes, cnmf_obj=cnmf_obj,k=3)
```

```{r}
xeno_usage = py$usage_by_calc
```

# programs expression {.tabset}
```{r echo=TRUE, fig.height=7, fig.width=9, results='asis'}

names (xeno_usage) = c("Hypoxia","TNFa","Cell_cycle")
#add each metagene to metadata
for (i  in 1:ncol(xeno_usage)) {
  metage_metadata = xeno_usage %>% dplyr::select(i)
  xeno = AddMetaData(object = xeno,metadata = metage_metadata,col.name = names(xeno_usage)[i])
}

print_tab(plt = FeaturePlot(object = xeno,features = colnames(xeno_usage)),title = "umap expression")


```



# programs regulation {.tabset}
```{r echo=TRUE,  results='asis'}
metagenes_mean_compare(dataset = xeno,time.point_var = "treatment",prefix = "model",patient.ident_var = "orig.ident",pre_on = c("NT","OSI"),test = "wilcox.test",programs = c("Hypoxia","TNFa","Cell_cycle"))

```

```{r}
DotPlot(object = xeno, features =  c("Hypoxia","TNFa","Cell_cycle"),group.by  = 'treatment')
```



# CC signature  {.tabset}


```{r results='asis'}
hallmark_name = "HALLMARK_G2M_CHECKPOINT"
annotation = plot_genes_cor(dataset = xeno,geneIds = genesets[[hallmark_name]],height = 3)
```

```{r results='asis'}
hallmark_name = "HALLMARK_G2M_CHECKPOINT"
annotation = plot_genes_cor(dataset = xeno,hallmark_name = hallmark_name,num_of_clusters = 4)
```

```{r results='asis'}
#choose clusters
chosen_clusters = c(1,2)
print (paste("chosen_clusters = ", chosen_clusters %>% toString()))

#UMAP expression of signature
chosen_genes = annotation[["myannotation"]] %>% dplyr::filter(cluster %in% chosen_clusters) %>% rownames() #take relevant genes
score <- apply(xeno@assays$RNA@data[chosen_genes,],2,mean)
xeno=AddMetaData(xeno,score,hallmark_name)
scoresAndIndices <- getPathwayScores(xeno@assays$RNA@data, chosen_genes)
xeno=AddMetaData(xeno,scoresAndIndices$pathwayScores,paste0(hallmark_name,"_sipsic"))
print_tab(FeaturePlot(object = xeno, features = paste0(hallmark_name,"_sipsic")),title = "Expression")
print_tab(FeaturePlot(object = xeno, features = hallmark_name),title = "Expression")
```

```{r echo=TRUE, results='asis'}

#plot signature distribution
cc_scores = FetchData(object = xeno,vars = "HALLMARK_G2M_CHECKPOINT")
plt  =  ggplot(cc_scores, aes(x=HALLMARK_G2M_CHECKPOINT)) +
  geom_density()+
   geom_vline(
    aes(xintercept=mean(cc_scores$HALLMARK_G2M_CHECKPOINT) + sd(cc_scores$HALLMARK_G2M_CHECKPOINT) ,color="1 SD"),
          linetype="dashed", size=1)+
    geom_vline(
    aes(xintercept=mean(cc_scores$HALLMARK_G2M_CHECKPOINT) + 2*sd(cc_scores$HALLMARK_G2M_CHECKPOINT) ,color="2 SD"),
          linetype="dashed", size=1)

print_tab(plt = plt,title = "dist")

# Plot assignment 

cc_scores = cc_scores %>% mutate(is_cycling = if_else(condition = 
                                          HALLMARK_G2M_CHECKPOINT > mean(HALLMARK_G2M_CHECKPOINT) 
                                          +sd(HALLMARK_G2M_CHECKPOINT),
                                        true = "cycling",
                                        false = "non_cycling"))
xeno = AddMetaData(object = xeno,metadata = cc_scores$is_cycling,col.name = "is_cycling")
print_tab(plt = DimPlot(object = xeno,group.by = "is_cycling") , title = "assignment 1 sd")


# cc_scores = cc_scores %>% mutate(is_cycling = if_else(condition = 
#                                           HALLMARK_G2M_CHECKPOINT > mean(HALLMARK_G2M_CHECKPOINT) + 2*sd(cc_scores$HALLMARK_G2M_CHECKPOINT),
#                                         true = "cycling",
#                                         false = "non_cycling"))
# xeno = AddMetaData(object = xeno,metadata = cc_scores$is_cycling,col.name = "is_cycling")
# print_tab(plt = DimPlot(object = xeno,group.by = "is_cycling") , title = "assignment 2 sd")


```




```{r echo=TRUE, results='asis'}
 df  = FetchData(object = xeno,vars = c("is_cycling","treatment")) %>% 
    filter (treatment %in% c("NT","OSI")) %>% 
    droplevels() 
  test = fisher.test(table(df))
    
  library(ggstatsplot)

    plt = ggbarstats(
    df, is_cycling, treatment,
    results.subtitle = FALSE,
    subtitle = paste0(
      "Fisher's exact test", ", p-value = ",
       round(test$p.value,13))
    )
  
print_tab(plt = plt,title = "fisher")
```
## correlation to nmf program
```{r}
cor_res = cor(xeno$Cell_cycle,xeno[[hallmark_name]])
print(paste("correlation of TNFa program to", hallmark_name,":", cor_res))

cor_res = cor(xeno$Cell_cycle,xeno[[paste0(hallmark_name,"_sipsic")]])
print(paste("correlation of Cell_cycle program to", paste0(hallmark_name,"_sipsic"),":", cor_res))
```


# TNFa signature  {.tabset}

```{r results='asis'}
hallmark_name = "HALLMARK_TNFA_SIGNALING_VIA_NFKB"
annotation = plot_genes_cor(dataset = xeno,hallmark_name = hallmark_name,num_of_clusters = 6)


```


```{r echo=TRUE, results='asis'}
chosen_clusters = c(1,5)
print (paste("chosen_clusters = ", chosen_clusters %>% toString()))

#UMAP expression of signature
chosen_genes = annotation[["myannotation"]] %>% dplyr::filter(cluster %in% chosen_clusters) %>% rownames() #take relevant genes
score <- apply(xeno@assays$RNA@data[chosen_genes,],2,mean)
xeno=AddMetaData(xeno,score,hallmark_name)
scoresAndIndices <- getPathwayScores(xeno@assays$RNA@data, chosen_genes)
xeno=AddMetaData(xeno,scoresAndIndices$pathwayScores,paste0(hallmark_name,"_sipsic"))
print_tab(FeaturePlot(object = xeno, features = paste0(hallmark_name,"_sipsic")),title = "Expression")
print_tab(FeaturePlot(object = xeno, features = hallmark_name),title = "Expression")
```
## correlation to nmf program

```{r}
cor_res = cor(xeno$TNFa,xeno[[hallmark_name]])
print(paste("correlation of TNFa program to", hallmark_name,":", cor_res))

cor_res = cor(xeno$TNFa,xeno[[paste0(hallmark_name,"_sipsic")]])
print(paste("correlation of TNFa program to", paste0(hallmark_name,"_sipsic"),":", cor_res))
```

```{r echo=TRUE, results='asis'}
cc_scores = FetchData(object = xeno,vars = hallmark_name)

plt  =  ggplot(cc_scores, aes(x=HALLMARK_TNFA_SIGNALING_VIA_NFKB)) +
  geom_density()+
   geom_vline(
    aes(xintercept=mean(HALLMARK_TNFA_SIGNALING_VIA_NFKB) + sd(HALLMARK_TNFA_SIGNALING_VIA_NFKB) ,color="1 SD"),
          linetype="dashed", size=1)+
    geom_vline(
    aes(xintercept=mean(HALLMARK_TNFA_SIGNALING_VIA_NFKB) + 2*sd(HALLMARK_TNFA_SIGNALING_VIA_NFKB) ,color="2 SD"),
          linetype="dashed", size=1)+
    geom_vline(
    aes(xintercept=mean(HALLMARK_TNFA_SIGNALING_VIA_NFKB) ,color="mean"),
          linetype="dashed", size=1)

print_tab(plt = plt,title = "dist")
```



# weighted ks test   {.tabset}
```{r ,results='asis'}

  
ranked_list1 = gep_scores %>% select(1) %>% as_vector() %>% setNames(rownames(gep_scores)) %>% sort(decreasing = TRUE) 
ranked_list2 = gep_scores %>% select(2) %>% as_vector() %>% setNames(rownames(gep_scores)) %>% sort(decreasing = TRUE) 
ranked_list3 = gep_scores %>% select(3) %>% as_vector() %>% setNames(rownames(gep_scores)) %>% sort(decreasing = TRUE) 



ranked_list = list("gep1" = ranked_list1, "gep2" = ranked_list2,"gep3" = ranked_list3)
hyp_obj <- hypeR(ranked_list2, genesets_env, test="kstest", fdr=0.05, plotting=F,background = rownames(gep_scores))
plots = hyp_dots(hyp_obj,size_by = "none")
print_tab(plt = plots[[1]],title = "GEP 1")
print_tab(plt = plots[[2]],title = "GEP 2")
print_tab(plt = plots[[3]],title = "GEP 3")

```

```{r}
mhyp_obj <- fgsea.wrapper(ranked_list, genesets)
hyp_dots(mhyp_obj)
```

# APC signature  {.tabset}

```{r results='asis'}
hallmark_name = "KEGG_ANTIGEN_PROCESSING_AND_PRESENTATION"
annotation = plot_genes_cor(dataset = xeno,hallmark_name = hallmark_name,num_of_clusters = 4,canocical = T)
```


```{r echo=TRUE, results='asis'}
chosen_clusters = c(1,3)
print (paste("chosen_clusters = ", chosen_clusters %>% toString()))

#UMAP expression of signature
chosen_genes = annotation[["myannotation"]] %>% dplyr::filter(cluster %in% chosen_clusters) %>% rownames() #take relevant genes
score <- apply(xeno@assays$RNA@data[chosen_genes,],2,mean)
xeno=AddMetaData(xeno,score,hallmark_name)
scoresAndIndices <- getPathwayScores(xeno@assays$RNA@data, chosen_genes)
xeno=AddMetaData(xeno,scoresAndIndices$pathwayScores,paste0(hallmark_name,"_sipsic"))
print_tab(FeaturePlot(object = xeno, features = paste0(hallmark_name,"_sipsic")),title = "Expression")
print_tab(FeaturePlot(object = xeno, features = hallmark_name),title = "Expression")
```

## correlatino to nmf
```{r}
cor_res = cor(xeno$TNFa,xeno[[hallmark_name]])
print(paste("correlation of TNFa program to", hallmark_name,":", cor_res))

cor_res = cor(xeno$TNFa,xeno[[paste0(hallmark_name,"_sipsic")]])
print(paste("correlation of TNFa program to", paste0(hallmark_name,"_sipsic"),":", cor_res))
```

# Signatures regulation {.tabset}
```{r echo=TRUE, results='asis'}
metagenes_mean_compare(dataset = xeno,time.point_var = "treatment",prefix = "model",patient.ident_var = "orig.ident",pre_on = c("NT","OSI","res"),programs = c("TNFa","HALLMARK_TNFA_SIGNALING_VIA_NFKB","KEGG_ANTIGEN_PROCESSING_AND_PRESENTATION"))

```
   
# intersection {.tabset}

```{r results='asis'}
top_ot = gep_scores3[order(gep_scores3[,2],decreasing = T),2,drop = F]%>% head(200) %>% rownames()
hallmark_name = "KEGG_ANTIGEN_PROCESSING_AND_PRESENTATION"
geneIds= genesets[[hallmark_name]]
chosen_genes = intersect(geneIds,top_ot)
print(chosen_genes)
score <- apply(xeno@assays$RNA@data[chosen_genes,],2,mean)
xeno=AddMetaData(xeno,score,paste0(hallmark_name,"intersected"))
scoresAndIndices <- getPathwayScores(xeno@assays$RNA@data, chosen_genes)
xeno=AddMetaData(xeno,scoresAndIndices$pathwayScores,paste0(hallmark_name,"_intersected_sipsic"))
print_tab(FeaturePlot(object = xeno, features = paste0(hallmark_name,"_intersected_sipsic")),title = "Expression")
print_tab(FeaturePlot(object = xeno, features = paste0(hallmark_name,"intersected")),title = "Expression")
```

```{r}
cor_res = cor(xeno$TNFa,xeno[[paste0(hallmark_name,"intersected")]])
print(paste("correlation of TNFa program to", paste0(hallmark_name,"intersected"),":", cor_res))

```

```{r}
top_ot = gep_scores3[order(gep_scores3[,2],decreasing = T),2,drop = F]%>% head(200) %>% rownames()
hallmark_name = "HALLMARK_TNFA_SIGNALING_VIA_NFKB"
geneIds= genesets[[hallmark_name]]
chosen_genes = intersect(geneIds,top_ot)
print(chosen_genes)
score <- apply(xeno@assays$RNA@data[chosen_genes,],2,mean)
xeno=AddMetaData(xeno,score,paste0(hallmark_name,"intersected"))
scoresAndIndices <- getPathwayScores(xeno@assays$RNA@data, chosen_genes)
xeno=AddMetaData(xeno,scoresAndIndices$pathwayScores,paste0(hallmark_name,"_intersected_sipsic"))
print_tab(FeaturePlot(object = xeno, features = paste0(hallmark_name,"_intersected_sipsic")),title = "Expression")
print_tab(FeaturePlot(object = xeno, features = paste0(hallmark_name,"intersected")),title = "Expression")
```
```{r}
cor_res = cor(xeno$TNFa,xeno[[paste0(hallmark_name,"intersected")]])
print(paste("correlation of TNFa program to", paste0(hallmark_name,"intersected"),":", cor_res))

```
#  Norm pathways  {.tabset}
```{r fig.width=10}
sum_scores= xeno$HIF_targets+xeno$KEGG_ANTIGEN_PROCESSING_AND_PRESENTATION+xeno$HALLMARK_G2M_CHECKPOINT

new_score = xeno$KEGG_ANTIGEN_PROCESSING_AND_PRESENTATION/ sum_scores
xeno %<>% AddMetaData(metadata = new_score,col.name = "KEGG_ANTIGEN_PROCESSING_AND_PRESENTATION_norm")

new_score = (xeno$HIF_targets/ sum_scores)%>% replace(is.na(.), 0)
xeno %<>% AddMetaData(metadata = new_score,col.name = "HIF_targets_norm")

new_score = xeno$HALLMARK_G2M_CHECKPOINT/ sum_scores
xeno %<>% AddMetaData(metadata = new_score,col.name = "HALLMARK_G2M_norm")
FeaturePlot(object = xeno,features = c("HALLMARK_G2M_norm", "KEGG_ANTIGEN_PROCESSING_AND_PRESENTATION_norm","HIF_targets_norm"))
```

```{r echo=TRUE, results='asis'}
metagenes_mean_compare(dataset = xeno,time.point_var = "treatment",prefix = "model",patient.ident_var = "orig.ident",pre_on = c("NT","OSI","res"),programs = c("TNFa","HALLMARK_TNFA_SIGNALING_VIA_NFKB","KEGG_ANTIGEN_PROCESSING_AND_PRESENTATION","KEGG_ANTIGEN_PROCESSING_AND_PRESENTATION_norm"))

```

# program 2 clustering
```{r}
num_of_clusters = 7
annotation = plot_genes_cor(dataset = xeno,hallmark_name = NULL,num_of_clusters = num_of_clusters,geneIds = top_ot)
```

# all clusters expression {.tabset}
```{r results='asis',fig.width=14}
for (chosen_clusters in 1:num_of_clusters) {
  chosen_genes = annotation[["myannotation"]] %>% dplyr::filter(cluster == chosen_clusters) %>% rownames() #take relevant genes
  # print(chosen_genes)
  hyp_obj <- hypeR(chosen_genes, genesets_env, test = "hypergeometric", fdr=1, plotting=F,background = rownames(gep_scores))

   scoresAndIndices <- getPathwayScores(xeno@assays$RNA@data, chosen_genes)
  xeno=AddMetaData(xeno,scoresAndIndices$pathwayScores,paste0("cluster",chosen_clusters))

  
  print_tab(plt = 
              hyp_dots(hyp_obj,size_by = "none",title = paste0("cluster",chosen_clusters))+
              FeaturePlot(object = xeno,features = paste0("cluster",chosen_clusters)),
            title = chosen_clusters)
  
  cor_res = cor(xeno$TNFa,xeno[[paste0("cluster",chosen_clusters)]])
# print(paste("correlation of TNFa program to", paste0("cluster",chosen_clusters),":", cor_res))


}


```
# correlation of all clusters
```{r ,fig.width=14}
for (chosen_clusters in 1:num_of_clusters) {
  chosen_genes = annotation[["myannotation"]] %>% dplyr::filter(cluster == chosen_clusters) %>% rownames() #take relevant genes
  cor_res = cor(xeno$TNFa,xeno[[paste0("cluster",chosen_clusters)]])
  print(paste("correlation of TNFa program to", paste0("cluster",chosen_clusters),":", cor_res))

}


```
# clusters 4 and 5 
```{r fig.width=14}
chosen_clusters = c(4,5)
  chosen_genes = annotation[["myannotation"]] %>% dplyr::filter(cluster %in% chosen_clusters) %>% rownames() #take relevant genes
  
  hyp_obj <- hypeR(chosen_genes, genesets_env, test = "hypergeometric", fdr=1, plotting=F,background = rownames(gep_scores))

   scoresAndIndices <- getPathwayScores(xeno@assays$RNA@data, chosen_genes)
   score_name = paste0("cluster_", paste0(chosen_clusters, collapse = "_"))
  xeno=AddMetaData(xeno,scoresAndIndices$pathwayScores,score_name)

  hyp_dots(hyp_obj,size_by = "none")+
    FeaturePlot(object = xeno,features = score_name)

  cor_res = cor(xeno$TNFa,xeno[[score_name]])
print(paste("correlation of TNFa program to", score_name,":", cor_res))
```




# Title {.tabset}


```{r results='asis'}
library(biomaRt)
ensembl = useEnsembl(biomart="ensembl", dataset="hsapiens_gene_ensembl")

for (chosen_clusters in 1:num_of_clusters) {
  chosen_genes = annotation[["myannotation"]] %>% dplyr::filter(cluster == chosen_clusters) %>% rownames() #take relevant genes
  genedesc <- getBM(attributes=c('external_gene_name','wikigene_description',"interpro_description"), filters = 'external_gene_name', values = chosen_genes, mart =ensembl)  
  print_tab(genedesc,title = chosen_clusters)
}


```

# Programs regulation {.tabset}
```{r echo=TRUE, results='asis'}
metagenes_mean_compare(dataset = xeno,time.point_var = "treatment",prefix = "model",patient.ident_var = "orig.ident",pre_on = c("NT","OSI","res"),programs = c("TNFa","HALLMARK_TNFA_SIGNALING_VIA_NFKB","KEGG_ANTIGEN_PROCESSING_AND_PRESENTATION","KEGG_ANTIGEN_PROCESSING_AND_PRESENTATIONintersected","cluster_4_5","cluster7"))

```

# HIF_targets- Hypoxia correlation  {.tabset}
```{r echo=TRUE, fig.height=8, fig.width=6, results='asis'}
# plot correlation for every subset of hif targets
for (genes in list(hif_targets,xeno_cluster_3_genes,xeno_cluster_3_2_genes)) {
  hif_targets_by_tp = FetchData(object = xeno,vars = c(genes)) %>% rowSums() %>% as.data.frame() #mean expression
  hif_targets_by_tp[,2] = xeno$Hypoxia
  
  names(hif_targets_by_tp) = c("hif_targets","hypoxia_program")
  
  
  
  p1 = ggplot(hif_targets_by_tp, aes(x=hif_targets, y=hypoxia_program)) + 
      geom_point()+
    geom_density_2d(aes(color = ..level..)) +
    geom_smooth(method=lm) +
    stat_cor(method = "pearson", label.x = 20, label.y = 1.1)+
    scale_color_viridis_c()
  
  p2 = ggplot(hif_targets_by_tp, aes(x=hif_targets, y=hypoxia_program)) + 
    geom_bin2d() +
    theme_bw()+ scale_fill_gradientn(limits=c(0,1100), breaks=seq(0, 1100, by=200), colours=c("blue","yellow","red"))+ 
    stat_cor(method = "pearson", label.x = 20, label.y = 1.1)+
    geom_smooth(method=lm) 
  
  p = ggarrange(plotlist = list(p1,p2),nrow  = 2)  
  
  print_tab(plt = p,title = "geom_bin2d")
}


```

# UMAPS
```{r fig.height=7, fig.width=10}
hif_targets_by_tp = FetchData(object = xeno,vars = c(hif_targets)) %>% rowSums() %>% as.data.frame() #mean expression
hif_targets_by_tp[,2] = xeno$Hypoxia
names(hif_targets_by_tp) = c("hif_targets","hypoxia_program")

high_hif_low_hypoxia_cells = hif_targets_by_tp %>% filter(hif_targets>25 & hypoxia_program < 0.2) %>% rownames()
low_hif_high_hypoxia_cells = hif_targets_by_tp %>% filter(hif_targets<15 & hypoxia_program > 0.6) %>% rownames()

hif_targets_by_tp = FetchData(object = xeno,vars = c(hif_targets)) %>% rowSums() %>% as.data.frame() #mean expression
xeno = AddMetaData(object = xeno, metadata = hif_targets_by_tp,col.name = "HIF_targets_score")
cells_to_highlight =  list(high_hif_low_hypoxia_cells = high_hif_low_hypoxia_cells, low_hif_high_hypoxia_cells = low_hif_high_hypoxia_cells)

DimPlot(object = xeno, cells.highlight = cells_to_highlight, cols.highlight = c("red","blue"), cols = "gray", order = TRUE)
FeaturePlot(object = xeno,features = c( "HIF_targets_score","Hypoxia","Cell_cycle" ))

```
```{r fig.height=8, fig.width=10}
hif_hypoxia_correlated_cells = colnames(xeno) [!colnames(xeno) %in% high_hif_low_hypoxia_cells & !colnames(xeno) %in% high_hif_low_hypoxia_cells]
hif_targets_score = FetchData(object = xeno,vars = c(hif_targets)) %>% rowSums() %>% as.data.frame() #mean expression
hif_targets_score[,2] = xeno$Hypoxia
names(hif_targets_score) = c("hif_targets","hypoxia_program")
hif_targets_score  = hif_targets_score %>% mutate(type = case_when(hif_targets>25 & hypoxia_program < 0.2 ~ "high_hif_low_hypoxia",
                                                                   hif_targets<15 & hypoxia_program > 0.6 ~ "low_hif_high_hypoxia",
                                                                   hif_targets>25 & hypoxia_program > 0.6 ~ "high_hif_high_hypoxia",
                                                                   hif_targets<15 & hypoxia_program <0.2 ~ "high_hif_high_hypoxia",
                                                                   TRUE ~ "other"))
xeno = AddMetaData(object = xeno,metadata = hif_targets_score[,"type", drop = F],col.name = "score_correlation")
xeno = SetIdent(object = xeno,value = "score_correlation")
DimPlot(object = xeno,group.by = "score_correlation")
markers = FindMarkers(object = xeno, ident.1 = "high_hif_low_hypoxia",ident.2 = "high_hif_high_hypoxia",densify = T)
```
```{r echo=TRUE, fig.height=8, fig.width=6, results='asis'}
updeg = markers %>% filter(p_val_adj<0.05 & avg_log2FC>0) %>% rownames()

new_hif_targets = hif_targets[!hif_targets %in% updeg]
hif_targets_by_tp = FetchData(object = xeno,vars = c(new_hif_targets)) %>% rowSums() %>% as.data.frame() #mean expression
  hif_targets_by_tp[,2] = xeno$Hypoxia
  
  names(hif_targets_by_tp) = c("hif_targets","hypoxia_program")
  
  
  
  p1 = ggplot(hif_targets_by_tp, aes(x=hif_targets, y=hypoxia_program)) + 
      geom_point()+
    geom_density_2d(aes(color = ..level..)) +
    geom_smooth(method=lm) +
    stat_cor(method = "pearson", label.x = 20, label.y = 1.1)+
    scale_color_viridis_c()
  
  p2 = ggplot(hif_targets_by_tp, aes(x=hif_targets, y=hypoxia_program)) + 
    geom_bin2d() +
    theme_bw()+ scale_fill_gradientn(limits=c(0,1100), breaks=seq(0, 1100, by=200), colours=c("blue","yellow","red"))+ 
    stat_cor(method = "pearson", label.x = 20, label.y = 1.1)+
    geom_smooth(method=lm) 
  
  p = ggarrange(plotlist = list(p1,p2),nrow  = 2)  
  
  print_tab(plt = p,title = "geom_bin2d")


```


```{r}
upreg_hif_targets = hif_targets[hif_targets %in% updeg]
upreg_hif_targets_expr = FetchData(object = xeno,vars = c(upreg_hif_targets)) %>% rowSums() %>% as.data.frame() #mean expression
xeno = AddMetaData(object = xeno, metadata = upreg_hif_targets_expr,col.name = "upreg_hif_targets_score")
FeaturePlot(object = xeno,features = "upreg_hif_targets_score")
```


# Calculate usage without cc in sum to 1
```{python}
import numpy as np
import scanpy as sc
xeno_expression = r.xeno_expression
xeno_vargenes = r.xeno_vargenes
tpm =  compute_tpm(xeno_expression)
usage_by_calc = get_usage_from_score(counts=xeno_expression,tpm=tpm,genes=xeno_vargenes, cnmf_obj=cnmf_obj,k=3,sumTo1=False)
```

```{r}
all_metagenes_noSumTo1 = py$usage_by_calc
tnf_and_hypoxia = all_metagenes_noSumTo1[,1:2]
tnf_and_hypoxia = apply(X = tnf_and_hypoxia, MARGIN = 1, sum_2_one) %>% t() %>%  as.data.frame()
tnf_and_hypoxia[is.na(tnf_and_hypoxia)] <- 0 #replace NAN's with 0.
```

```{r echo=TRUE, fig.height=8, fig.width=6, results='asis'}
# plot correlation for every subset of hif targets
for (genes in list(hif_targets,xeno_cluster_3_genes,xeno_cluster_3_2_genes)) {
  hif_targets_by_tp = FetchData(object = xeno,vars = c(genes)) %>% rowSums() %>% as.data.frame() #mean expression
  hif_targets_by_tp[,2] = tnf_and_hypoxia[,1]
  # hif_targets_by_tp[,2] = xeno$Hypoxia
  
  names(hif_targets_by_tp) = c("hif_targets","hypoxia_program")
  
  
  
  p1 = ggplot(hif_targets_by_tp, aes(x=hif_targets, y=hypoxia_program)) + 
      geom_point()+
    geom_density_2d(aes(color = ..level..)) +
    geom_smooth(method=lm) +
    stat_cor(method = "pearson", label.x = 20, label.y = 1.1)+
    scale_color_viridis_c()
  
  p2 = ggplot(hif_targets_by_tp, aes(x=hif_targets, y=hypoxia_program)) + 
    geom_bin2d() +
    theme_bw()+ scale_fill_gradientn(limits=c(0,1100), breaks=seq(0, 1100, by=200), colours=c("blue","yellow","red"))+ 
    stat_cor(method = "pearson", label.x = 20, label.y = 1.1)+
    geom_smooth(method=lm) 
  
  p = ggarrange(plotlist = list(p1,p2),nrow  = 2)  
  
  print_tab(plt = p,title = "geom_bin2d")
}


```

# UMAPS
```{r fig.height=6, fig.width=8}
hif_targets_by_tp = FetchData(object = xeno,vars = c(hif_targets)) %>% rowSums() %>% as.data.frame() #mean expression
hif_targets_by_tp[,2] = tnf_and_hypoxia[,1]
names(hif_targets_by_tp) = c("hif_targets","hypoxia_program")

high_hif_low_hypoxia_cells = hif_targets_by_tp %>% filter(hif_targets>25 & hypoxia_program < 0.2) %>% rownames()

hif_targets_by_tp = FetchData(object = xeno,vars = c(hif_targets)) %>% rowSums() %>% as.data.frame() #mean expression
xeno = AddMetaData(object = xeno, metadata = hif_targets_by_tp,col.name = "HIF_targets_score")
xeno = AddMetaData(object = xeno, metadata = tnf_and_hypoxia[,1],col.name = "Hypoxia2")

DimPlot(object = xeno, cells.highlight = high_hif_low_hypoxia_cells, cols.highlight = "red", cols = "gray", order = TRUE)
FeaturePlot(object = xeno,features = c( "HIF_targets_score","Hypoxia2","Cell_cycle" ))
FeaturePlot(object = xeno,features = c("Hypoxia2"))
DimPlot(object = xeno,group.by = "orig.ident")

```

# Hypoxia raw
```{r fig.height=8, fig.width=10}
xeno = AddMetaData(object = xeno,metadata = all_metagenes_noSumTo1[,1],col.name = "hypoxia_raw")
FeaturePlot(object = xeno,features = "hypoxia_raw") + scale_color_gradientn(colours = rainbow(5), limits = c(0, 3000))
```



<script src="https://hypothes.is/embed.js" async></script>
